---
title: "Course project 1"
author: "Zimu"
date: "7/25/2021"
output: 
  html_document:
    keep_md: true
---

```{r, echo=TRUE, results=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

# Loading and preprocessing the data

```{r, echo = TRUE }
activity = read.csv("activity.csv")
activity$date = as.Date(activity$date)
```


# What is mean total number of steps taken per day?

## Calculate the total number of steps taken per day
## Histogram of the total number of steps taken each day

```{r, echo = TRUE }
step_date <- activity %>% group_by(date) %>% summarize(steps=sum(steps))
ggplot(step_date, aes(x=steps)) + geom_histogram()
```

## What is mean total number of steps taken per day?
Mean and median number of steps taken each day:

```{r, echo = TRUE }
mean_step <- mean(step_date$steps)
median_step <- median(step_date$steps)
```

Mean step: `r mean_step`
Median step: `r median_step`

## Time series plot of the average number of steps taken
```{r, echo = TRUE }
step_interval <- activity %>% group_by(interval) %>% summarize(steps=mean(steps,na.rm=TRUE))
ggplot(step_interval,aes(x=interval, y=steps)) + geom_line()
max_interval = step_interval[which.max(step_interval$steps),1]
```

The `r max_interval`  5-minute interval contains the maximum number of steps.

Max steps: `r max(step_interval$steps)`

# Imputing missing values

## Total number of missing values in the dataset

A total of `r sum(is.na(activity))` values are missing in the dataset. 

## Missing values are imputed using the mean value of the same 5 minute interval across different dates

```{r, echo = TRUE }
activity_imputed <- activity
for (i in 1:length(activity_imputed$steps)){
  if(is.na(activity_imputed[i,"steps"])){
    cor_interval <- activity_imputed[i,"interval"]
    imputed_steps <- mean(activity[activity$interval==cor_interval,"steps"],na.rm=TRUE)
    activity_imputed[i,"steps"] <- as.integer(imputed_steps)
  }
}
```

## Make a histogram of the total number of steps taken each day.

```{r, echo = TRUE }
step_date_imputed <- activity_imputed %>% group_by(date) %>% summarize(steps=sum(steps,na.rm=TRUE))
ggplot(step_date_imputed, aes(x=steps)) + geom_histogram()
```

## Calculate and report the mean and median total number of steps taken per day. 
```{r, echo = TRUE }
mean_step_imputed <- mean(step_date_imputed$steps)
median_step_imputed <- median(step_date_imputed$steps)
```

Mean step with imputed values: `r mean_step_imputed`
Median step with imputed values: `r median_step_imputed`

# Are there differences in activity patterns between weekdays and weekends?
```{r, echo = TRUE }
activity_imputed <- activity_imputed %>% mutate(weekday = weekdays(as.Date(activity_imputed$date)))
activity_imputed <- activity_imputed %>% mutate(weekend = weekday %in% c("Saturday","Sunday"))
activity_imputed$weekend <- if_else(activity_imputed$weekend,"Weekend","Weekday")
activity_imputed$weekend <- factor(activity_imputed$weekend)
steps_weekend <- activity_imputed %>% group_by(weekend, interval) %>% summarize(steps=mean(steps))
ggplot(steps_weekend,aes(x=interval, y=steps)) + geom_line() + facet_grid(weekend~.)
```